\LARGE \bf Enhancing Security in Multi-Robot Systems through Co-Observation Planning, Reachability Analysis, and Network Flow

### 2025-05-31 response_2.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Academic Paper Response Letter. 
% LaTeX Template
% Version 1.0 (20/07/2023)
% 
% This template refers to the toolbox tutorial (in Chinese):
% https://liam.page/2016/07/22/using-the-tcolorbox-package-to-create-a-new-theorem-environment/
% Please refer to the tutorial for your customization. 
% 
% Template license: Creative Commons CC BY 4.0
%
% Author: Yuhan Xie
% Contact email: yhxie0107@gmail.com
% Any revision advice is welcome!
% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{article}
%% Language and font encodings
\input{preamble/common}
\usepackage[english]{babel}
\usepackage{times} % make the section title in Times New Roman

\usepackage{booktabs}
\usepackage{tabu}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[a4paper,top=2.5cm,bottom=2cm,left=1.7cm,right=1.7cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
% \usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{indentfirst}
\usepackage{graphicx}
% \usepackage{subfigure}
\usepackage{float}
\usepackage{threeparttable} 
\usepackage{multirow}
\usepackage{url}
\usepackage{xcolor}
\pagenumbering{arabic}

\usepackage[most]{tcolorbox}
\newcommand{\rrtstar}{$\texttt{RRT}^\texttt{*}$}
\newcommand{\SubItem}[1]{
    {\setlength\itemindent{15pt} \item[-] #1}
}

\newtheorem{constraint}{ADMM constraint}
\newtheorem{example}{Example}
\newcommand{\re}{\tcblower \underline{\textbf{Response:}}\quad}
\newcommand{\rv}{{\large{\underline{\textbf{Revision:}}}}\quad}
\usepackage[export]{adjustbox} 
\usepackage[capitalise]{cleveref}
\usepackage{xcolor}
\newcommand{\new}[1]{\textcolor{blue}{#1}}
\newcommand{\news}{\color{blue}}
\graphicspath{{Figs/}}

\begin{document}

\newtcbtheorem[auto counter, number within = section]{cmt}{Comment}{
	colbacktitle = black!60!white, colframe = black!60!white,
	colback = black!5!white,
	fonttitle=\bfseries,%fontupper=\itshape,
}{t}

\noindent
Dear Editor and Reviewers:

\vspace{0.2cm}
\noindent
We would like to express our sincere gratitude to the Editor-in-Chief, Associate Editor, and anonymous reviewers for their time and constructive feedback throughout the review process. Following the acceptance of our manuscript, we have carefully addressed the remaining minor issues related to typos, formatting, and clarity. All necessary corrections have been made, and the updated changes are highlighted in \new{blue} for clarity. We appreciate the opportunity to further improve the manuscript and are confident that these refinements contribute to a more polished and readable final version.

\vspace{0.2cm}
\noindent
Yours sincerely,

\noindent
Ziqi Yang, Roberto Tron

\vspace{1cm}
\begin{cmt*}{}{}
Some minor formatting adjustments have been made to ensure that the overall content remains within the page limit. These changes are purely presentational and do not affect the technical content of the paper.
\end{cmt*}
\rv 
\textbf{Appendix B}
\setcounter{proposition}{2}
\begin{proposition}\label{prop:HProperty}
 {\news The matrix $H$ has the following properties:
  \begin{enumerate}
  \item\label{it:rotation} It is a rotation, i.e. (a) $H\transpose H=I$; (b) $\det(H)=1$
  % \begin{enumerate}
  %   \item\label{it:orthonormality} $H\transpose H=I$;
  %   \item\label{it:determinant} $\det(H)=1$.
  %   \end{enumerate}
  \item\label{it:transformation} $\nu_2=H \nu_1$.
  \end{enumerate}}
\end{proposition}
\begin{proof}
 {\news For subclaim~\ref{it:rotation}a,  since $u\transpose u=1$:
  \begin{equation}
    H\transpose H=H^2=4uu\transpose u u\transpose - 4uu\transpose +I^2=I.
  \end{equation}}
\end{proof}
\newpage
\section{Response to reviewer 1}
\begin{cmt*}{}{}
We thank the reviewer for their time and constructive feedback during the review process. We are pleased that no further comments were raised and appreciate their role in helping improve the quality of the manuscript.
\end{cmt*}
\vspace{0.2cm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Rewiewer 1
\section{Response to reviewer 2} 
\begin{cmt*}{}{}
We sincerely thank the reviewer for their constructive comments during the initial review, which significantly contributed to improving the quality of this paper. We also appreciate the reviewer’s acknowledgment of the improvements made in the previous revision. In this final round, we have carefully addressed the additional comments raised and made the necessary corrections in the manuscript. Our detailed responses are provided below.
\end{cmt*}
\vspace{0.1cm}
%%%%%%%%%%%%% Comment 1.0
\begin{cmt}{}{}        %% use * when do not need enumerate
1. There still are some typos, eg.,

   a) in Fig 4, the legend is Region 1 and Region 2, but in the text,
you use Zone1 and Zone2.

   b) Sec 2D, first paragraph, $q^{\epsilon} \in F^{\epsilon}$?
	\re All typos mentioned have been fixed.
\end{cmt}
\rv 
\textbf{Section II.D}

\newcommand{\oFE}{o}
To use the definition of reachability ellipsoid from the section above as computational constraints for ADMM, we first apply a differentiable rigid body transformation to reposition the ellipse $\cE$ from the global frame $\cF$ to a canonical frame $\cF_\cE$, where the origin of $\cF_\cE$ is at the ellipsoid's center $\oFE = \frac{1}{2}(q_1+q_2)$, and the first axis of $\cF_\cE$ is aligned with the foci (see \cref{fig:EllipseConstraintExample} for an illustration). 
From this definition, a unit vector along the first axis of the ellipsoid in the frames $\cF$ and $\cF_\cE$ is given by $\nu_\cF = \frac{q_2-q_1}{\norm{q_2-q_1}}$ and $\nu_\cE =[1,0,0]^T$, respectively.
The full transformation from coordinates \new{$q^\cE\in\cF_\cE$} to $q\in\cF$, and its inverse, are then parametrized by a rotation $R^\cF_\cE$ and a translation $o^\cF_\cE$ as (we drop the subscript and superscript from $R^\cF_\cE$ and $o^\cF_\cE$ to simplify the notation):\setcounter{figure}{3}
\begin{figure}[h]
  \centering
  \subfloat[Grid world result\label{fig:Grid-example-application}]{\includegraphics[height = 0.3\linewidth,trim =0cm 0cm 0cm 0cm, clip,valign=t]{Grid_world}}
  \subfloat[Secured result \label{fig:ReachabilitySimulation}]{\includegraphics[height=0.3\linewidth,valign=t]{FinalResult-v2}}
  \caption{ Trajectory design of a three-robot system based on the grid world result. \new{Region} 1 is an obstacle, \new{region} 2 and \new{region} 3 are forbidden regions. (\ref{fig:Grid-example-application}) Result of the APMAPF algorithm in $8\times8$ grid-world without the map exploring objective. (\ref{fig:ReachabilitySimulation}) The secured trajectory with the incorporation of co-observation generated by APMAPF and additional reachability constraints. Attack attempts into the forbidden region (red dashed line) will cause team 1 miss the next scheduled co-observation.}
  \label{fig:example-application}
\end{figure}

\textbf{Section II.H}

The simulation result, shown in Fig.~\ref{fig:ReachabilitySimulation}, displays reachability regions as black ellipsoids, demonstrating empty intersections with \new{region} 2 and 3. Explicit constraints between reachability regions and obstacles are not activated, assuming basic obstacle avoidance capabilities in robots. The intersections between obstacles and ellipsoids, as observed between agent 3 and \new{region} 1, are deemed tolerable. All constraints are satisfied, and agents have effectively spread across the map for optimal exploration tasks. 
\begin{cmt}{}{}
	2. In algorithm 1, what is the reason for using $t_0 + 1$ and $t_2 - 1$?

	3. In Fig 5, no explanation is made in the caption
\re The algorithm aims to find the furthest waypoint such that the reachability region between waypoint at $t_0$ and a later waypoint that does not intersect with any forbidden region. Since evaluating the reachability region between the same waypoint (e.g., $t_0$ to $t_0$) is not meaningful, we exclude such cases and begin the search from $t_0 +1$ The same logic applies in the reverse direction, where we stop at $t_2 -1$ instead of including $t_2$. For better readability, we reorder the sequence of the variables ($v_0$ to $v_3$) and add additional indicator in the illustration figure (Fig.5).
\end{cmt}
\rv
\setcounter{figure}{4}
\begin{algorithm}
	\caption{Checkpoint Graph Initialization for Team $\cI_p$}\label{alg:graph}
  \begin{algorithmic}[1]
	{\news \State \textbf{Initialization:} Set $t_{0} \leftarrow 0$ and $t_{3} \leftarrow T$, and add $(q_{p}(t_0),t_0)$ and $(q_{p}(t_3),t_3)$ to $V_{p}$ as \emph{start} and \emph{end vertices}.
	\While{$\mathcal{E}(q_{p}(t_{0}),q_{p}(t_3),t_{0},t_3) \cap F \neq \emptyset$ \textbf{and} $t_{0} \leq t_3$}
		\State \textbf{Forward Search:} Find the largest $t_{1} \in \{t_{0}+1, \dots, t_3\}$ such that $\mathcal{E}(q_{p}(t_{0}),q_{p}(t_{1}), t_{0},t_{1}) \cap F = \emptyset$. Once found, add $(q_{p}(t_{1}),t_{1})$ to $V_{p}$.
		\State \textbf{Backward Search:} Find the smallest $t_2\in\{t_{1}, \dots, t_3-1\}$ such that $
		\mathcal{E}(q_{p}(t_3),q_{p}(t_2), t_3,t_2) \cap F = \emptyset.$ Once found, add $(q_{p}(t_2),t_2)$ to $V_{p}$.
		\State \textbf{Update Indices:} Set $t_{0} \leftarrow t_{1}$ and $t_3 \leftarrow t_2$.
	\EndWhile}
	\end{algorithmic}
	\end{algorithm}
\begin{figure}[h]
	\centering
    \subfloat{\includegraphics[width=0.47\linewidth,valign=c]{SC_gen_1_v3}}
    \subfloat{\includegraphics[width=0.47\linewidth,valign=c]{SC_gen_2_v3}}
    \caption{\news Example of \cref{alg:graph}. Start at both ends, each trajectory finds the largest reachability region avoiding forbidden areas, then repeats the process on remaining segments to build secure checkpoints.}\label{fig:checkpoint-generate}
\end{figure}

\end{document}

### 2025-05-31 paper.tex
Accepted version

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We address security challenges in multi-robot systems (MRS) where adversaries may take control of compromised robots to gain unauthorized physical access to forbidden regions. We propose a novel multi-robot optimal planning algorithm that integrates mutual observations and introduces {\news \emph{reachability constraints} and \emph{co-observations} between robots to enhance security}. Our formulation can guarantee that, even with adversarial movements, compromised robots cannot breach forbidden regions without missing scheduled co-observations. \new{We use ellipsoidal over-approximations to reachability regions for efficient intersection checking and gradient computation. Furthermore, to enhance system resilience and tackle feasibility challenges, we introduce the notion of \emph{sub-teams}; these cohesive units replace individual robots along each trajectory; we plan \emph{cross-trajectory co-observation} that use redundant robots to switch between different trajectories, securing multiple sub-teams without requiring modifications to the plans. We formulate cross-trajectories plans by solving a \emph{network-flow vertex path coverage problem} on the \emph{checkpoint graph} generated from the original unsecured MRS trajectories, providing the same security guarantees against plan-deviation attacks. We demonstrate the effectiveness and robustness of our proposed algorithm through simulations, which significantly strengthen the security of multi-robot systems in the face of adversarial threats.}
\end{abstract}
\begin{IEEEkeywords}
  Multi-robot system, cyber-physical attack, trajectory optimization, reachability analysis, network flow
\end{IEEEkeywords}

\section{Introduction}\label{sec:introduction}
Multi-robot systems (MRS) have found wide applications in various fields. \new{Although offering numerous advantages, the distributed nature and dependence on network communication make the MRS vulnerable to cyber threats, such as unauthorized access, malicious attacks, and data manipulation \cite{brunner2010infiltrating}.} This paper addresses a specific scenario in which robots are compromised by \new{physically masquerading} attackers and directed toward \emph{forbidden regions}, which may contain security-sensitive equipment or human workers. {\news A countermeasure to these \emph{plan-deviation attacks} \cite{wardega2019resilience, wardega2023byzantine, wardega2023hola} is to use the onboard sensing capabilities of the robots themselves to perform inter-robot \emph{co-observations} and detect unusual behavior. These mutual observations establish a \emph{co-observation schedule} alongside the planned path, ensuring that any attempts by a compromised robot to violate safety constraints (such as transgressing into forbidden regions) would break the observation plan and be promptly detected.}

A preliminary version of this approach is presented in \cite{yang2020multi,yang2021multi}. That work extended the grid-world solution from \cite{wardega2019resilience} to continuous configuration space, and incorporated the co-observation planning as constraints in \new{the alternating direction method of multipliers (ADMM)-based trajectory optimization solver to accommodate for more flexible objectives. However, such plans are not guaranteed to exist, and the secured trajectory always comes at the cost of overall system performance. As an extension of the previous works \cite{yang2020multi,yang2021multi}, we introduce reachability analysis and network flow to the security problem to find a balance between security and performance.}

In this paper, we incorporate additional reachability analysis during the planning phase, incorporating constraints based on sets of locations that agents could potentially reach, referred to as \emph{reachability regions}, as a novel perspective to the existing literature. This approach enforces an empty intersection between forbidden regions and the reachability region during trajectory optimization, preventing undetected attacks if an adversary gains control of the robots. We utilize an ellipsoidal boundary to constrain the search space and formulate the ellipsoid as the reachability region constraint. We present a mathematical formulation of the reachability regions \new{as spatio-temporal constraints} compatible with the solver in \cite{yang2020multi}.

\new{Furthermore, we address the feasibility and optimality challenges. For a MRS with unsecured optimal trajectories (without security constraints), we introduce redundant robots and form them into \emph{sub-teams} with the original ones.} These redundant robots are assigned to establish additional co-observations, termed \emph{cross-trajectory co-observations}, within and across different sub-teams. The proposed algorithm focuses on the movement plans for the additional robots, distinct from those dedicated to task objectives, indicating when they should stay with their current sub-team and when they should deviate to join another. This strategy allows sub-teams to preserve the optimal unsecured trajectories (as illustrated in \cref{fig:cross-traj-comparison-set}) without requiring the entire \emph{sub-team} to maintain close proximity to other teams during co-observation events. The cross-trajectory co-observation problem is transformed as a \new{Multi-Agent Pathfinding (MAPF)} problem on roadmap (represented as directed graph) and solved as a network flow problem \cite{yu2013multi}.

@@

    \caption{{\news (\ref{fig:example-plan-dev}) The attacker deviates the red robot into a forbidden region.(\ref{fig:example-co-observation}) Red and blue robots are scheduled to co-observe each other during the task, resulting in a new secured trajectory (solid lines) replacing the optimal ones (dashed line). (\ref{fig:example-cross-traj}) The blue team sends one robot to observe the red team (solid blue line) while having the rest of the robots follow the optimal trajectory.}\label{fig:cross-traj-comparison-set}}
\end{figure}

@@

Graph-based search is another extensively explored approach in trajectory planning problems, such as formation control \cite{tanner2004leader,hu2019distributed} and MAPF problems \cite{stern2019multi}. In traditional MAPF formulations, environments are abstracted as graphs, with nodes representing positions and edges denoting possible transitions between positions, and solved as a network flow problem \cite{yu2013multi,yu2016optimal}. This formulation allows for the application of combinatorial network flow algorithms and linear program techniques, offering efficient and more flexible solutions to the planning problem. %In contrast with such approach, our research in \cref{sec:cross-trajectory} abstracts the existing Multi-Robot System (MRS) trajectory (instead of the environment) into a directed graph and formulates the cross-trajectory co-observation paths as flows.
% \rtron{Cite the paper from the reviewer, as well as the paper from Tedrake on the IRIS decomposition, saying that ellipsoids have been used to create approximations of the free configuration space, not to bound reacheability, and not for security application here.}

% Reachability analysis is a important step in the security and safety verification synthesis for cyber-physical systems (CPSs) \cite{gueguen2009safety,ding2020secure}. One typical approach involves computing an over-approximation of the reachable space for checking safety properities. Previous studies \cite{kurzhanski2000ellipsoidal,lakhal2019interval,maiga2015comprehensive} often use geometric representation, such as zonotopes and ellipsoids, as a compact enclosure of the reachability sets. In the realm of online safety assessment against cyber attacks, \cite{kwon2017reachability} computed the reachable set of CPS states achievable by potential cyber attacks and compared it with the safe region, set based on current state estimation and environmental conditions. 
\new{Reachability analysis is essential for security and safety verification in cyber-physical systems (CPSs) \cite{gueguen2009safety, ding2020secure}, often involving an over-approximation of reachable space to verify safety properties. Geometric methods like zonotopes and ellipsoids are commonly used to enclose reachability sets compactly \cite{kurzhanski2000ellipsoidal, lakhal2019interval, maiga2015comprehensive}. For online safety assessments against cyber attacks, \cite{kwon2017reachability} compute reachable CPS states under attacks and compared them with a safe region based on state estimation. By incorporating an additional security measure that provides two secured states, the reachability region in our work is over-approximated using ellipsoids. This is inspired by the ellipsoidal \emph{heuristic sampling domain} in \cite{gammell2014informed} for the $\mathtt{RRT^*}$ algorithm to simplify the sampling region between start and goal locations. Ellipsoids are also used in other path-planning methods like Iterative Regional Inflation by Semidefinite programming (IRIS) \cite{deits2015computing, ray2022free} and the Safe Flight Corridor (SFC) \cite{liu2017planning, fan2024flying}. Differing from the reachability region that requires mapping all reachable states given several known states, the ellipsoids of IRIS and SFC focus on approximating the safe collision-free space rather than addressing security applications.}

@@

{\news The rest of this paper is organized as follows. \cref{sec:ADMM-planning} introduces the ADMM-based optimal trajectory solver and the security constraints. \cref{sec:cross-trajectory} introduces the enhanced security planning algorithm through cross-trajectory co-observation. \cref{sec:summary} concludes this article.

\noindent\myparagraph{Notation} 
In this paper, we use non-bold symbols to denote single-agent states (e.g. $q_{ij}$) and scalars, and bold symbols to represent aggregated states of single robot and multiple robots (e.g. $\vq$). }

\section{Secured Multi-robot trajectory planning}\label{sec:ADMM-planning}
We formulate the planning problem as an optimal trajectory optimization problem to minimize arbitrary smooth objective functions. {\news We denote the trajectory as $\vq_i = [q_{i0}\dots q_{iT}]\in\real{m\times T}$, where $q_{ij}\in\real{m}$ is the waypoint of agent $i$ in a $m$ dimensional state space, $T$ is the time horizon. For a total of $n_p$ robots, trajectories can be represented as an aggregated vector $\vq = \stack(\vq_1,\dots,\vq_{n_p})\in \real{m n_p\times T}$, where $\stack(\cdot)$ denotes the vertical stacking operation. }
\new{The overall goal is to minimize or maximize an objective function $\varPhi(\vq):\real{n m T} \to \real{}$ under a set of nonlinear constraints described by a set $\Omega$, which is given by the intersection of spatio-temporal sets given by traditional path planning constraints and the security constraints (co-observation schedule, reachability analysis).} Formally:
\begin{equation}\label{eq:general-problem}
	\begin{split}
		\min/\max & \quad \varPhi(\vq)\\
		\textrm{subject to} &\quad \vq \in \Omega.
	\end{split}
\end{equation}
To give a concrete example of the cost $\varPhi$ and the set $\Omega$, we introduce a representative application that will be used for all the simulations throughout the paper.

\begin{example}\label{example:map_exploration}
\new{Robots are tasked with navigating an unknown task space to collect sensory data and reconstruct a field (as illustrated in~\cref{fig:illustration}). The task space is modeled as a grid, where each grid point has an associated value tracked by a Kalman Filter (KF) \cite{anderson2012optimal}. The KF estimates uncertainty at each point through its covariance $P_j$ updated based on the measurements taken by robots along the trajectory $\vq$. Measurement quality, modeled by a Gaussian radial basis function, decreases with distance from the robot. The optimization objective $\varPhi(\vq)=\max_j P_j(\vq)$ is to minimize the maximum uncertainty in reconstructing the field (detailed in \cite{yang2020multi}). }
\end{example}
\begin{figure}
  \centering
  \subfloat[\new{Map exploration task}\label{fig:illustration}]{\includegraphics[width = 0.45\linewidth,, trim = 0.3cm 1.3cm 0.3cm 2cm,valign=c]{illustration figure_1}} 
  \subfloat[Continuous world trajectory \label{fig:SecurityBreak}]{\includegraphics[width=0.45\linewidth, trim = 0.3cm 1.3cm 0.3cm 2cm, valign=c]{SecurityBreak_v3}}
  % \subfloat[\new{Grid world result.}\label{fig:Grid-example-application}]{\includegraphics[width = 0.45\linewidth]{Grid_world}}
  \caption{\new{(\ref{fig:illustration}) Illustration of a 3 robot map exploration task case. (\ref{fig:SecurityBreak}) The unsecured trajectory design optimized for a map exploration task. Potential security breaches, indicated by red dashed lines, highlight paths that could allow unauthorized access to forbidden regions.}}
\end{figure}

{\news Ideally, robots are expected to spread out in a Boustrophedon pattern to efficiently explore the task space, optimize coverage, and minimize reconstruction uncertainty (\cref{fig:SecurityBreak}). However, we consider the possibility of threats following the \emph{physical masquerade attack} model \cite{wardega2019resilience}, where a compromised insider (robot) masquerading as a properly functioning agent, attempts to gain access to unauthorized locations without detection. The attacker leverages full knowledge of the motion plan and exploits the compromised robot to provide false self-reports to the central entity (CE). These malicious deviations remain undetected as long as observations from uncompromised robots align with the motion plan. The corresponding security requirement can then be formally defined as:}
\begin{definition}\label{def:secured-plan}
  A multi-robot trajectory plan is \new{secured against} plan-deviation attacks if it \new{ensures} that any potential deviations to these forbidden regions will cause the corresponding robot to miss their next co-observation with other robots. 
\end{definition}

\subsection{Preliminaries}
\subsubsection{Differentials}
\new{We define the differential of a map $f(x):\real{d_1}\to\real{d_2}$ at a point $x_0$ as the unique matrix $\partial_x f \in\real{d_1\times d_2}$ such that}
\begin{equation}\label{equ:dt_to_dx}
  \left.\dert f\bigl(x(t)\bigr)\right|_{t=0}=\partial_x f\bigl(x(0)\bigr) \dot{x}(0),
\end{equation}
where $t\mapsto x(t)\in\real{d_1}$ is a smooth parametric curve such that $x(0)=x$ with any arbitrary tangent $\dot{x}(0)$. For brevity, we will use $\dot f$ for $\dert f$ and $\partial_x f$ for $\frac{\partial f}{\partial x}$. 
%The differentials $\partial_x f$ is derived through \eqref{equ:dt_to_dx} having $\dot f$ divided by $\dot x$.

\new{With a slight abuse of notation, we use the same notation $\partial_xf$ for the differential of a matrix-valued function with scalar arguments $f:\real{d_1 \times d_3}\to\real{d_2}$.  Note that in this case \eqref{equ:dt_to_dx} is still formally correct, although the RHS needs to be interpreted as applying $\partial_x f$ as a linear operator to $\dot{x}$.}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We address security challenges in multi-robot systems (MRS) where adversaries may take control of compromised robots to gain unauthorized physical access to forbidden regions. We propose a novel multi-robot optimal planning algorithm that integrates mutual observations and introduces \emph{reachability constraints} and \emph{co-observations} between robots to enhance security. Our formulation can guarantee that, even with adversarial movements, compromised robots cannot breach forbidden regions without missing scheduled co-observations. We use ellipsoidal over-approximations to reachability regions for efficient intersection checking and gradient computation. Furthermore, to enhance system resilience and tackle feasibility challenges, we introduce the notion of \emph{sub-teams}; these cohesive units replace individual robots along each trajectory; we plan \emph{cross-trajectory co-observation} that use redundant robots to switch between different trajectories, securing multiple sub-teams without requiring modifications to the plans. We formulate cross-trajectories plans by solving a \emph{network-flow vertex path coverage problem} on the \emph{checkpoint graph} generated from the original unsecured MRS trajectories, providing the same security guarantees against plan-deviation attacks. We demonstrate the effectiveness and robustness of our proposed algorithm through simulations, which significantly strengthen the security of multi-robot systems in the face of adversarial threats.
\end{abstract}
\begin{IEEEkeywords}
  Multi-robot system, cyber-physical attack, trajectory optimization, reachability analysis, network flow
\end{IEEEkeywords}

\section{Introduction}\label{sec:introduction}
Multi-robot systems (MRS) have found wide applications in various fields. Although offering numerous advantages, the distributed nature and dependence on network communication make the MRS vulnerable to cyber threats, such as unauthorized access, malicious attacks, and data manipulation \cite{brunner2010infiltrating}. This paper addresses a specific scenario in which robots are compromised by physically masquerading attackers and directed toward \emph{forbidden regions}, which may contain security-sensitive equipment or human workers. A countermeasure to these \emph{plan-deviation attacks} \cite{wardega2019resilience, wardega2023byzantine, wardega2023hola} is to use the onboard sensing capabilities of the robots themselves to perform inter-robot \emph{co-observations} and detect unusual behavior. These mutual observations establish a \emph{co-observation schedule} alongside the planned path, ensuring that any attempts by a compromised robot to violate safety constraints (such as transgressing into forbidden regions) would break the observation plan and be promptly detected.

A preliminary version of this approach is presented in \cite{yang2020multi,yang2021multi}. That work extended the grid-world solution from \cite{wardega2019resilience} to continuous configuration space, and incorporated the co-observation planning as constraints in the alternating direction method of multipliers (ADMM)-based trajectory optimization solver to accommodate for more flexible objectives. However, such plans are not guaranteed to exist, and the secured trajectory always comes at the cost of overall system performance. As an extension of the previous works \cite{yang2020multi,yang2021multi}, we introduce reachability analysis and network flow to the security problem to find a balance between security and performance.

In this paper, we incorporate additional reachability analysis during the planning phase, incorporating constraints based on sets of locations that agents could potentially reach, referred to as \emph{reachability regions}, as a novel perspective to the existing literature. This approach enforces an empty intersection between forbidden regions and the reachability region during trajectory optimization, preventing undetected attacks if an adversary gains control of the robots. We utilize an ellipsoidal boundary to constrain the search space and formulate the ellipsoid as the reachability region constraint. We present a mathematical formulation of the reachability regions as spatio-temporal constraints compatible with the solver in \cite{yang2020multi}.

Furthermore, we address the feasibility and optimality challenges. For a MRS with unsecured optimal trajectories (without security constraints), we introduce redundant robots and form them into \emph{sub-teams} with the original ones. These redundant robots are assigned to establish additional co-observations, termed \emph{cross-trajectory co-observations}, within and across different sub-teams. The proposed algorithm focuses on the movement plans for the additional robots, distinct from those dedicated to task objectives, indicating when they should stay with their current sub-team and when they should deviate to join another. This strategy allows sub-teams to preserve the optimal unsecured trajectories (as illustrated in \cref{fig:cross-traj-comparison-set}) without requiring the entire \emph{sub-team} to maintain close proximity to other teams during co-observation events. The cross-trajectory co-observation problem is transformed as a Multi-Agent Pathfinding (MAPF) problem on roadmap (represented as directed graph) and solved as a network flow problem \cite{yu2013multi}.

@@

    \caption{ (\ref{fig:example-plan-dev}) The attacker deviates the red robot into a forbidden region.(\ref{fig:example-co-observation}) Red and blue robots are scheduled to co-observe each other during the task, resulting in a new secured trajectory (solid lines) replacing the optimal ones (dashed line). (\ref{fig:example-cross-traj}) The blue team sends one robot to observe the red team (solid blue line) while having the rest of the robots follow the optimal trajectory.}\label{fig:cross-traj-comparison-set}
\end{figure}

@@

Graph-based search is another extensively explored approach in trajectory planning problems, such as formation control \cite{tanner2004leader,hu2019distributed} and MAPF problems \cite{stern2019multi}. In traditional MAPF formulations, environments are abstracted as graphs, with nodes representing positions and edges denoting possible transitions between positions, and solved as a network flow problem \cite{yu2013multi,yu2016optimal}. This formulation allows for the application of combinatorial network flow algorithms and linear program techniques, offering efficient and more flexible solutions to the planning problem.

Reachability analysis is essential for security and safety verification in cyber-physical systems (CPSs) \cite{gueguen2009safety, ding2020secure}, often involving an over-approximation of reachable space to verify safety properties. Geometric methods like zonotopes and ellipsoids are commonly used to enclose reachability sets compactly \cite{kurzhanski2000ellipsoidal, lakhal2019interval, maiga2015comprehensive}. For online safety assessments against cyber attacks, \cite{kwon2017reachability} compute reachable CPS states under attacks and compared them with a safe region based on state estimation. By incorporating an additional security measure that provides two secured states, the reachability region in our work is over-approximated using ellipsoids. This is inspired by the ellipsoidal \emph{heuristic sampling domain} in \cite{gammell2014informed} for the $\mathtt{RRT^*}$ algorithm to simplify the sampling region between start and goal locations. Ellipsoids are also used in other path-planning methods like Iterative Regional Inflation by Semidefinite programming (IRIS) \cite{deits2015computing, ray2022free} and the Safe Flight Corridor (SFC) \cite{liu2017planning, fan2024flying}. Differing from the reachability region that requires mapping all reachable states given several known states, the ellipsoids of IRIS and SFC focus on approximating the safe collision-free space rather than addressing security applications.

@@

The rest of this paper is organized as follows. \cref{sec:ADMM-planning} introduces the ADMM-based optimal trajectory solver and the security constraints. \cref{sec:cross-trajectory} introduces the enhanced security planning algorithm through cross-trajectory co-observation. \cref{sec:summary} concludes this article.

\noindent\myparagraph{Notation} 
In this paper, we use non-bold symbols to denote single-agent states (e.g. $q_{ij}$) and scalars, and bold symbols to represent aggregated states of single robot and multiple robots (e.g. $\vq$).

\section{Secured Multi-robot trajectory planning}\label{sec:ADMM-planning}
We formulate the planning problem as an optimal trajectory optimization problem to minimize arbitrary smooth objective functions. We denote the trajectory as $\vq_i = [q_{i0}\dots q_{iT}]\in\real{m\times T}$, where $q_{ij}\in\real{m}$ is the waypoint of agent $i$ in a $m$ dimensional state space, $T$ is the time horizon. For a total of $n_p$ robots, trajectories can be represented as an aggregated vector $\vq = \stack(\vq_1,\dots,\vq_{n_p})\in \real{m n_p\times T}$, where $\stack(\cdot)$ denotes the vertical stacking operation. 
The overall goal is to minimize or maximize an objective function $\varPhi(\vq):\real{n m T} \to \real{}$ under a set of nonlinear constraints described by a set $\Omega$, which is given by the intersection of spatio-temporal sets given by traditional path planning constraints and the security constraints (co-observation schedule, reachability analysis). Formally:
\begin{equation}\label{eq:general-problem}
	\begin{split}
		\min/\max & \quad \varPhi(\vq)\\
		\textrm{subject to} &\quad \vq \in \Omega.
	\end{split}
\end{equation}
To give a concrete example of the cost $\varPhi$ and the set $\Omega$, we introduce a representative application that will be used for all the simulations throughout the paper.

\begin{example}\label{example:map_exploration}
Robots are tasked with navigating an unknown task space to collect sensory data and reconstruct a field (as illustrated in~\cref{fig:illustration}). The task space is modeled as a grid, where each grid point has an associated value tracked by a Kalman Filter (KF) \cite{anderson2012optimal}. The KF estimates uncertainty at each point through its covariance $P_j$ updated based on the measurements taken by robots along the trajectory $\vq$. Measurement quality, modeled by a Gaussian radial basis function, decreases with distance from the robot. The optimization objective $\varPhi(\vq)=\max_j P_j(\vq)$ is to minimize the maximum uncertainty in reconstructing the field (detailed in \cite{yang2020multi}). 
\end{example}
\begin{figure}
  \centering
  \subfloat[Map exploration task\label{fig:illustration}]{\includegraphics[width = 0.45\linewidth,, trim = 0.3cm 1.3cm 0.3cm 2cm,valign=c]{illustration figure_1}} 
  \subfloat[Continuous world trajectory \label{fig:SecurityBreak}]{\includegraphics[width=0.45\linewidth, trim = 0.3cm 1.3cm 0.3cm 2cm, valign=c]{SecurityBreak_v3}}
  \caption{(\ref{fig:illustration}) Illustration of a 3 robot map exploration task case. (\ref{fig:SecurityBreak}) The unsecured trajectory design optimized for a map exploration task. Potential security breaches, indicated by red dashed lines, highlight paths that could allow unauthorized access to forbidden regions.}
\end{figure}

 Ideally, robots are expected to spread out in a Boustrophedon pattern to efficiently explore the task space, optimize coverage, and minimize reconstruction uncertainty (\cref{fig:SecurityBreak}). However, we consider the possibility of threats following the \emph{physical masquerade attack} model \cite{wardega2019resilience}, where a compromised insider (robot) masquerading as a properly functioning agent, attempts to gain access to unauthorized locations without detection. The attacker leverages full knowledge of the motion plan and exploits the compromised robot to provide false self-reports to the central entity (CE). These malicious deviations remain undetected as long as observations from uncompromised robots align with the motion plan. The corresponding security requirement can then be formally defined as:
\begin{definition}\label{def:secured-plan}
  A multi-robot trajectory plan is secured against plan-deviation attacks if it ensures that any potential deviations to these forbidden regions will cause the corresponding robot to miss their next co-observation with other robots. 
\end{definition}

\subsection{Preliminaries}
\subsubsection{Differentials}
We define the differential of a map $f(x):\real{d_1}\to\real{d_2}$ at a point $x_0$ as the unique matrix $\partial_x f \in\real{d_1\times d_2}$ such that
\begin{equation}\label{equ:dt_to_dx}
  \left.\dert f\bigl(x(t)\bigr)\right|_{t=0}=\partial_x f\bigl(x(0)\bigr) \dot{x}(0),
\end{equation}
where $t\mapsto x(t)\in\real{d_1}$ is a smooth parametric curve such that $x(0)=x$ with any arbitrary tangent $\dot{x}(0)$. For brevity, we will use $\dot f$ for $\dert f$ and $\partial_x f$ for $\frac{\partial f}{\partial x}$.

With a slight abuse of notation, we use the same notation $\partial_xf$ for the differential of a matrix-valued function with scalar arguments $f:\real{d_1 \times d_3}\to\real{d_2}$.  Note that in this case \eqref{equ:dt_to_dx} is still formally correct, although the RHS needs to be interpreted as applying $\partial_x f$ as a linear operator to $\dot{x}$.
